Toward theories of deliberated preferences

### 2020-12-07 simple.tex
Started abstract attempt

A supplementary problem might happen in some situations. Once an argumentative path has been presented to the individual, this cannot necessarily be undone easily: she might remember some of the arguments, and they may have a lasting influence on her perspective. This may render difficult to test several unrelated paths on a given individual: such tests require to wait long enough that it can be reasonably thought that the individual is no more under influence of the arguments he has seen. It is therefore desirable to point out which testing protocols permit to avoid these difficulties.
\commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}
An illustration of some consequences of these possibilities on our running example may be useful. 
\begin{example}[Hypothesis of memory]
	 The definition of $\ileadsto$ as given so far implicitly assumes that a given path can be tested at most once with a given individual, as it does not indicate what to conclude when a path $\hist$ leads $i$ to accept a proposition $\prop$ only some of the time. This is justified under the hypothesis that individuals remember arguments forever. 
	It is interesting to note that even in that case, instability of $i$ towards some set of argument can be observed. As an illustration of such an instability, recall \cref{ex:incompl}, which assumes that for some $\hist, \hist' \in \allhist$, $\hist \ileadsto P ≥ D$ and $\hist' \ileadsto D > P$.
	Consider $\hist = \left<\ar_0, \ar_1\right>$ and $\hist' = \left<\ar_0, \ar_1, \ar_0\right>$ and assume we tested first $\hist$ then $\hist'$ on the same individual. Even assuming that the individual remembers, when exposed to $\hist'$, having seen these two arguments during the previous test, she might be fundamentally undecided as to whether $P ≥ D$ or $D > P$, and pick items randomly when exposed to these two arguments in any order. As $\hist'$ is a sequence that starts with $\hist$, it is possible to test these two paths on $i$, even under the hypothesis of unbounded memory, and observe an instability of $i$ towards the arguments $\ar_0$ and $\ar_1$.

	Under the more realistic hypothesis that the individual forgets over time, not only can we test unrelated paths on a given individual, but we can test several times the same path. We must then define $\ileadsto$ differently. Define $\hist \ileadsto \prop$ iff the individual would be lead to accept $\prop$ on every presentation of $\hist$. It is then possible that $\exists \hist \in \allhist \suchthat \hist \nileadsto P ≥ D \land \hist \nileadsto D > P$.
	Note that this makes it impossible to observe that $\hist \ileadsto \prop$. This is not a problem for the methodology presented here: it is still possible to observe that $\hist \nileadsto \prop$, as it suffices to observe that the individual did not accept $\prop$ under at least one presentation of $\hist$. The falsification sequences defined here below rest only on that kind of observations.
\end{example}

Importantly, theories of \ac{DJ} must also be able to argue differently for different individuals. Even considering individuals with the same \ac{DJ}, different individuals might consider as decisive different arguments. Accepting non trivial propositions often requires to accept several sub-propositions, that together permit to be convinced of the acceptability of the proposition. Depending on which of these sub-propositions the individual already considers to be acceptable, parts of the argumentation can be omitted, and, perhaps, must be omitted: an overly long argument, stating many things that the listener already knows, has fewer chance to convince the listener than a short argument that goes straight to the point about which the listener doubts. 
%Mathematicians do not write their proofs starting from the axioms of set theory: they use results that they assume the reader already is aware of, and are ready (in principle) to supply the proofs of these intermediate results if a reader has doubt about some of them. 
In other words, what is needed for a theory of \ac{DJ} is not just a capacity to produce a unique decisive argument, with the claim that this argument will convince individuals in $I$, but rather a capacity to produce dynamically arguments that depend on which counter-arguments the individual finds appealing, with a claim that this dynamic argumentation strategy will succeed in convincing individuals in $I$.

Achieving practical knowledge of \acp{DJ} is tackled by introducing descriptive theories of \ac{DJ}. They are defined in \cref{sec:descrdef}. An empirical theory for a given normative theory defines a subset of propositions of the agenda, that it claims are contained in the \acp{DJ} of individuals concerned by the normative theory. They are called the proposition that the theory supports. \commentYM{dans la mesure où, à ce stade, je ne comprends pas bien ce que contient une "normative theory", la séparation avec l' "empirical theory" n'est pas claire pour moi} 
%The theory also claims that it is able to argue in favor of the proposition it supports in such a way that individuals will be convinced of the quality of these arguments. This is accomplished by three binary relations, part of the definition of empiricial theories: one represents its initial arguments, one plans the possible attacks, and one defines its counter-arguments. An empirical theory also defines a concatenation operation able to dynamically build arguments from fragments of arguments.
%The theory claims that when one of its arguments in support of one of its claims is unconvincing to an individual, it is because of one of the anticipated attacks; and the theory claims to be able to provide counter-arguments to such attacks as they come, that will in turn be either convincing to the individual, or fail to be because of some anticipated attack. The theory is also constrained to have “finite depth”, meaning that it must set an a priori limit to the number of such iterations of attacks and counter-arguments.

%Thanks to these operators, an empirical theory may argue in a way that depends on the counter-arguments each individual finds appealing
%\commentYM{je ne comprends pas cette phrase, et en particulier le sens qu'il faut donner au terme "adjust"}. Instead of having to give a single decisive argument in support of its claims, the theory may react to exactly the counter-arguments that a given individual thinks relevant.

Such theories achieve what has been announced, namely, their conclusions rest solely on a weak normative axiom and on empirically falsifiable claims. 
Granted normative adequacy, if no falsification sequence exist for a given theory of \ac{DJ}, then it is valid, meaning that the propositions that it claims are \acp{DJ} are indeed \acp{DJ}.
\Cref{sec:validity} defines formally the notion of falsification sequence and proves this central claim.

This result permits to ensure that, under Normative adequacy, Validity is an empirical property. But it does not give a concrete way of arbitrating between conflicting theories. A supplementary important result, described in \cref{sec:compare}, is that when two theories disagree about incompatible propositions, one of the theories at least must fail some given sequence of empirical tests, which the proof describes.

These arbitration tests can all be performed on a single individual and are robust against the hypothesis of memory, which suggest that arbitration between theories could be performed reasonably easily.
By contrast, the falsification sequences used in the preceding result of \cref{sec:validity} may require to test unrelated paths on the same individual, which may be practically difficult, as acknowledged here above. I expect however that the second result, permitting to arbitrate between theories, will be more often used in practice, thus, that falsification tests of the previous kind will be comparatively rare. This article is content with the existence of a principled, systematic way of falsifying theories of \ac{DJ} and leaves to later studies the comparison of approaches from the practical point of view of easiness of falsifying them.

\subsection{Definition}
\label{sec:descrdef}
In this section, a normative theory of \ac{DJ} $(\allprops, \allargs, I, {\allleadsto})$ is supposed given.
A descriptive theory for a normative theory of \ac{DJ} is defined as a tuple $(\gleadsto, \dg)$, where ${\gleadsto} \subseteq \allargs × \allprops$ is an initial argumentation relation and $\dg$ is a defense relation; satisfying bounded depth.
These concepts are now defined formally.
\commentYM{puisque la théorie empirique est une théorie empirique D'UNE CERTAINE THEORIE NORMATIVE, je m'attendrais  à ce qu'on voit apparaître la théorie normative dans la formule de la théorie empirique, non ?}

>>>>

Achieving practical knowledge of \acp{DJ} is tackled by introducing descriptive theories of \ac{DJ}, the subject of the next section.

\subsection{Definition}
\label{sec:descrdef}
In this section, a normative theory of \ac{DJ} $(\allprops, \allargs, I, {\allleadsto})$ is supposed given.
A descriptive theory for a normative theory of \ac{DJ} is defined as a tuple $(\gC, \gPhi)$, where $\gC \subseteq \powerset{I × \allhist × \Phi}$ is a falsifiable claim and $\gPhi \subseteq \Phi$ are the proposition that the theory claims are in the \acp{DJ} of the individuals.
\commentYM{dans la mesure où, à ce stade, je ne comprends pas bien ce que contient une "normative theory", la séparation avec l' "empirical theory" n'est pas claire pour moi}

These concepts are now defined formally.
\commentYM{puisque la théorie empirique est une théorie empirique D'UNE CERTAINE THEORIE NORMATIVE, je m'attendrais  à ce qu'on voit apparaître la théorie normative dans la formule de la théorie empirique, non ?}

A claim $\gC \subseteq \powerset{I × \allhist × \Phi}$ represents the theory’s claim about the relation $\leadsto$: the theory claims that $\leadsto \in \gC$. 
A claim $\gC \subseteq \powerset{I × \allhist × \Phi}$ is falsifiable iff
